---
title: "Target 1 – GDP per capita growth"
output: pdf_document
---

First, we load all required packages.  

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(rnaturalearth)
library(rnaturalearthdata)
library(sf)
```

---

Here we import the three main datasets used for Target 1:
the continent classifications, GDP per capita from the World Bank,
and the NEET dataset (used later for context).

```{r load-data}
continents.according.to.our.world.in.data <- read.csv("~/Documents/GitHub/Data-Science-Coursework/continents-according-to-our-world-in-data.csv")
gdp.per.capita.worldbank <- read.csv("~/Documents/GitHub/Data-Science-Coursework/gdp-per-capita-worldbank.csv", header = FALSE)
youth.not.in.education.employment.training <- read.csv("~/Documents/GitHub/Data-Science-Coursework/youth-not-in-education-employment-training.csv")
```

---

This chunk cleans the continent data: we sort countries by continent,
remove Antarctica, and drop the `Year` column so it is easier to join
with the GDP dataset later.

```{r clean-continents}
Sorted_continents <- continents.according.to.our.world.in.data %>%
  arrange(Continent) %>%
  filter(Continent != "Antarctica")

Sorted_continents <- Sorted_continents[, -3]
```

---

We now clean the GDP per capita data: remove the first row (header row),
rename columns, and convert `Year` and the GDP variable to numeric.

```{r clean-gdp}
GDP_per_capita <- gdp.per.capita.worldbank[-1, ]

colnames(GDP_per_capita) <- c(
  "Entity",
  "Code",
  "Year",
  "GDP per capita, PPP (constant 2017 international $)"
)

GDP_per_capita <- GDP_per_capita %>%
  mutate(Year = as.numeric(Year)) %>%
  mutate(`GDP per capita, PPP (constant 2017 international $)` =
           as.numeric(`GDP per capita, PPP (constant 2017 international $)`))
```

---

Here we combine the continent information with the GDP dataset,
producing a single panel with GDP per capita, country, continent and year.

```{r merge-gdp-continents}
GDP_with_Continents <- full_join(Sorted_continents, GDP_per_capita)
```

---

This chunk creates six continent-specific GDP data frames,
each ordered by country and then by year.
They will be used to compute growth within each region.

```{r split-by-continent}
Africa_GDP <- GDP_with_Continents %>%
  filter(Continent == "Africa") %>%
  arrange(Entity, Year)

Asia_GDP <- GDP_with_Continents %>%
  filter(Continent == "Asia") %>%
  arrange(Entity, Year)

Europe_GDP <- GDP_with_Continents %>%
  filter(Continent == "Europe") %>%
  arrange(Entity, Year)

North_America_GDP <- GDP_with_Continents %>%
  filter(Continent == "North America") %>%
  arrange(Entity, Year)

Oceania_GDP <- GDP_with_Continents %>%
  filter(Continent == "Oceania") %>%
  arrange(Entity, Year)

South_America_GDP <- GDP_with_Continents %>%
  filter(Continent == "South America") %>%
  arrange(Entity, Year)
```

---

We define a helper function `Growth()` that, within each country,
computes GDP per capita growth (difference) and growth rate (percentage)
from one year to the next.
We then apply it to each continent.

```{r compute-growth}
Growth <- function(df) {
  df %>%
    group_by(Entity) %>%
    mutate(
      GDP_Growth = `GDP per capita, PPP (constant 2017 international $)` -
        lag(`GDP per capita, PPP (constant 2017 international $)`),
      GDP_Growth_Rate =
        (`GDP per capita, PPP (constant 2017 international $)` -
           lag(`GDP per capita, PPP (constant 2017 international $)`)) /
        lag(`GDP per capita, PPP (constant 2017 international $)`) * 100
    ) %>%
    ungroup()
}

Africa_GDP_Growth_Rate        <- Growth(Africa_GDP)
Asia_GDP_Growth_Rate          <- Growth(Asia_GDP)
Europe_GDP_Growth_Rate        <- Growth(Europe_GDP)
North_America_GDP_Growth_Rate <- Growth(North_America_GDP)
Oceania_GDP_Growth_Rate       <- Growth(Oceania_GDP)
South_America_GDP_Growth_Rate <- Growth(South_America_GDP)
```

---

Next, we define `Average_Growth()` which collapses a continent-level
data frame to average GDP growth rates per year (across its countries).
We then combine all continents into one dataset.

```{r continent-average-growth}
Average_Growth <- function(df) {
  df %>%
    select(Year, GDP_Growth_Rate, Continent) %>%
    group_by(Year, Continent) %>%
    summarise(Average_GDP_Growth_Rate = mean(GDP_Growth_Rate, na.rm = TRUE))
}

Continents_Growth_Rate <- rbind(
  Average_Growth(Africa_GDP_Growth_Rate),
  Average_Growth(Asia_GDP_Growth_Rate),
  Average_Growth(Europe_GDP_Growth_Rate),
  Average_Growth(North_America_GDP_Growth_Rate),
  Average_Growth(Oceania_GDP_Growth_Rate),
  Average_Growth(South_America_GDP_Growth_Rate)
)
```

---

This graph shows the average GDP per capita growth rate over time
for each continent using annual lines.

```{r plot-continent-growth, fig.width=8, fig.height=5}
Graph_Continents_Growth_Rate <- ggplot(
  Continents_Growth_Rate,
  aes(x = Year, y = Average_GDP_Growth_Rate, colour = Continent)
) +
  geom_line(linewidth = 0.2) +
  labs(
    x = "Year",
    y = "Average GDP per capita Growth Rate (%)",
    title = "Average GDP per capita Growth Rate against Time"
  ) +
  theme_minimal()

Graph_Continents_Growth_Rate
```

---

This alternative graph replaces annual lines with linear trends
(fitted via OLS) to better compare the long-run growth trajectories
of each continent.

```{r plot-continent-trends, fig.width=8, fig.height=5}
Graph_Continents_Average_Growth_Rate <- ggplot(
  Continents_Growth_Rate,
  aes(x = Year, y = Average_GDP_Growth_Rate, colour = Continent)
) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.2) +
  labs(
    x = "Year",
    y = "Average GDP per capita Growth Rate (%)",
    title = "Average GDP per capita Growth Rate Trend against Time"
  ) +
  theme_minimal()

Graph_Continents_Average_Growth_Rate
```

---

We now define plotting functions for country-level GDP growth
and its trend within each continent.
These will be used to generate detailed graphs by country.

```{r country-growth-functions}
Graph_Growth <- function(df, Continent) {
  ggplot(df, aes(x = Year, y = GDP_Growth, colour = Entity)) +
    geom_line(linewidth = 0.2) +
    labs(
      x = "Year",
      y = "GDP per capita Growth (PPP)",
      title = paste0("GDP per capita Growth against Time (", Continent, ")")
    ) +
    theme_minimal()
}

Graph_Average_Growth <- function(df, Continent) {
  ggplot(df, aes(x = Year, y = GDP_Growth, colour = Entity)) +
    geom_smooth(method = "lm", se = FALSE, linewidth = 0.2) +
    labs(
      x = "Year",
      y = "GDP per capita Growth (PPP)",
      title = paste0("GDP per capita Growth Trend against Time (", Continent, ")")
    ) +
    theme_minimal()
}
```

---

Here we actually call the functions above to produce
country-level GDP growth and trend graphs for each continent
(optional to print or save individually).

```{r country-growth-plots, eval=FALSE}
Graph_Growth(Africa_GDP_Growth_Rate,        Continent = "Africa")
Graph_Growth(Asia_GDP_Growth_Rate,          Continent = "Asia")
Graph_Growth(Europe_GDP_Growth_Rate,        Continent = "Europe")
Graph_Growth(North_America_GDP_Growth_Rate, Continent = "North America")
Graph_Growth(Oceania_GDP_Growth_Rate,       Continent = "Oceania")
Graph_Growth(South_America_GDP_Growth_Rate, Continent = "South America")

Graph_Average_Growth(Africa_GDP_Growth_Rate,        Continent = "Africa")
Graph_Average_Growth(Asia_GDP_Growth_Rate,          Continent = "Asia")
Graph_Average_Growth(Europe_GDP_Growth_Rate,        Continent = "Europe")
Graph_Average_Growth(North_America_GDP_Growth_Rate, Continent = "North America")
Graph_Average_Growth(Oceania_GDP_Growth_Rate,       Continent = "Oceania")
Graph_Average_Growth(South_America_GDP_Growth_Rate, Continent = "South America")
```

---

Similarly, we define functions to plot GDP growth **rates** (rather than levels)
for each country and to show smoothed trends.

```{r country-growth-rate-functions}
Graph_Growth_Rate <- function(df, Continent) {
  ggplot(df, aes(x = Year, y = GDP_Growth_Rate, colour = Entity)) +
    geom_line(linewidth = 0.2) +
    labs(
      x = "Year",
      y = "GDP per capita Growth Rate (%)",
      title = paste0("GDP per capita Growth Rate against Time (", Continent, ")")
    ) +
    theme_minimal()
}

Graph_Average_Growth_Rate <- function(df, Continent) {
  ggplot(df, aes(x = Year, y = GDP_Growth_Rate, colour = Entity)) +
    geom_smooth(method = "lm", se = FALSE, linewidth = 0.2) +
    labs(
      x = "Year",
      y = "GDP per capita Growth Rate (%)",
      title = paste0("GDP per capita Growth Rate Trend against Time (", Continent, ")")
    ) +
    theme_minimal()
}
```

---

This chunk illustrates how to generate those country-level
growth-rate plots for each continent (again, optional).

```{r country-growth-rate-plots, eval=FALSE}
Graph_Growth_Rate(Africa_GDP_Growth_Rate,        Continent = "Africa")
Graph_Growth_Rate(Asia_GDP_Growth_Rate,          Continent = "Asia")
Graph_Growth_Rate(Europe_GDP_Growth_Rate,        Continent = "Europe")
Graph_Growth_Rate(North_America_GDP_Growth_Rate, Continent = "North America")
Graph_Growth_Rate(Oceania_GDP_Growth_Rate,       Continent = "Oceania")
Graph_Growth_Rate(South_America_GDP_Growth_Rate, Continent = "South America")

Graph_Average_Growth_Rate(Africa_GDP_Growth_Rate,        Continent = "Africa")
Graph_Average_Growth_Rate(Asia_GDP_Growth_Rate,          Continent = "Asia")
Graph_Average_Growth_Rate(Europe_GDP_Growth_Rate,        Continent = "Europe")
Graph_Average_Growth_Rate(North_America_GDP_Growth_Rate, Continent = "North America")
Graph_Average_Growth_Rate(Oceania_GDP_Growth_Rate,       Continent = "Oceania")
Graph_Average_Growth_Rate(South_America_GDP_Growth_Rate, Continent = "South America")
```

---

## Part 1 – Income groups and growth targets

We first combine all continent-specific data into one large panel
of country-year observations with GDP growth and its rate.

```{r combine-all-countries}
Countries_GDP_Growth_Rate <- rbind(
  Africa_GDP_Growth_Rate,
  Asia_GDP_Growth_Rate,
  Europe_GDP_Growth_Rate,
  North_America_GDP_Growth_Rate,
  Oceania_GDP_Growth_Rate,
  South_America_GDP_Growth_Rate
)
```

---

Next we define income groups using GDP per capita thresholds,
restrict to the most recent 5-year window (2017–2021),
and ensure each country is assigned a single income group
based on its most recent value.

```{r income-groups}
Countries_Income_Group <- Countries_GDP_Growth_Rate %>%
  filter(between(Year, 2017, 2021)) %>%
  mutate(Income_Group = case_when(
    `GDP per capita, PPP (constant 2017 international $)` <= 1135   ~ "Low Income",
    `GDP per capita, PPP (constant 2017 international $)` <= 4495   ~ "Lower Middle Income",
    `GDP per capita, PPP (constant 2017 international $)` <= 13935  ~ "Upper Middle Income",
    `GDP per capita, PPP (constant 2017 international $)` > 13935   ~ "High Income"
  )) %>%
  group_by(Entity) %>%
  mutate(Income_Group = Income_Group[which.max(Year)]) %>%
  ungroup()
```

---

We then create four separate datasets, one for each income group.
These will be used both for individual country plots and for
continent-level averages within each income category.

```{r split-income-groups}
Low_Income <- Countries_Income_Group %>%
  filter(Income_Group == "Low Income")

Lower_Middle_Income <- Countries_Income_Group %>%
  filter(Income_Group == "Lower Middle Income")

Upper_Middle_Income <- Countries_Income_Group %>%
  filter(Income_Group == "Upper Middle Income")

High_Income <- Countries_Income_Group %>%
  filter(Income_Group == "High Income")
```

---

These four graphs show GDP per capita growth rates over time
for each income group, with a horizontal dashed line indicating
the corresponding SDG growth target.

```{r income-group-plots, fig.width=8, fig.height=5}
Graph_Low_Income <- ggplot(Low_Income) +
  aes(x = Year, y = GDP_Growth_Rate, colour = Entity) +
  geom_line(linewidth = 0.2) +
  geom_hline(yintercept = 7, linetype = "dashed", colour = "black", linewidth = 0.2) +
  labs(
    x = "Year",
    y = "GDP per capita Growth Rate (%)",
    title = "GDP per capita Growth Rate against Time (Low Income)"
  ) +
  theme_minimal()

Graph_Lower_Middle_Income <- ggplot(Lower_Middle_Income) +
  aes(x = Year, y = GDP_Growth_Rate, colour = Entity) +
  geom_line(linewidth = 0.2) +
  geom_hline(yintercept = 5, linetype = "dashed", colour = "black", linewidth = 0.2) +
  labs(
    x = "Year",
    y = "GDP per capita Growth Rate (%)",
    title = "GDP per capita Growth Rate against Time (Lower Middle Income)"
  ) +
  theme_minimal()

Graph_Upper_Middle_Income <- ggplot(Upper_Middle_Income) +
  aes(x = Year, y = GDP_Growth_Rate, colour = Entity) +
  geom_line(linewidth = 0.2) +
  geom_hline(yintercept = 4, linetype = "dashed", colour = "black", linewidth = 0.2) +
  labs(
    x = "Year",
    y = "GDP per capita Growth Rate (%)",
    title = "GDP per capita Growth Rate against Time (Upper Middle Income)"
  ) +
  theme_minimal()

Graph_High_Income <- ggplot(High_Income) +
  aes(x = Year, y = GDP_Growth_Rate, colour = Entity) +
  geom_line(linewidth = 0.2) +
  geom_hline(yintercept = 2, linetype = "dashed", colour = "black", linewidth = 0.2) +
  labs(
    x = "Year",
    y = "GDP per capita Growth Rate (%)",
    title = "GDP per capita Growth Rate against Time (High Income)"
  ) +
  theme_minimal()

Graph_Low_Income
```

*(The other three graphs can be printed the same way if needed.)*

---

We now compute mean GDP growth rates by continent and year
within each income group, to see how regions perform relative
to the target.

```{r average-growth-by-continent-income}
Average_Low_Income <- Low_Income %>%
  group_by(Continent, Year) %>%
  summarise(Mean_GDP_Growth_Rate = mean(GDP_Growth_Rate, na.rm = TRUE)) %>%
  ungroup()

Average_Lower_Middle_Income <- Lower_Middle_Income %>%
  group_by(Continent, Year) %>%
  summarise(Mean_GDP_Growth_Rate = mean(GDP_Growth_Rate, na.rm = TRUE)) %>%
  ungroup()

Average_Upper_Middle_Income <- Upper_Middle_Income %>%
  group_by(Continent, Year) %>%
  summarise(Mean_GDP_Growth_Rate = mean(GDP_Growth_Rate, na.rm = TRUE)) %>%
  ungroup()

Average_High_Income <- High_Income %>%
  group_by(Continent, Year) %>%
  summarise(Mean_GDP_Growth_Rate = mean(GDP_Growth_Rate, na.rm = TRUE)) %>%
  ungroup()
```

---

These four graphs show the continent-level mean growth rate for each
income group, again with horizontal dashed target lines.

```{r average-growth-by-continent-income-plots, fig.width=8, fig.height=5}
Graph_Average_Low_Income <- ggplot(Average_Low_Income) +
  aes(x = Year, y = Mean_GDP_Growth_Rate, colour = Continent) +
  geom_line(linewidth = 0.2) +
  geom_hline(yintercept = 7, linetype = "dashed", colour = "black") +
  labs(
    title = "Average GDP per capita Growth Rate per Continent against Time (Low Income)",
    x = "Year",
    y = "GDP per capita Growth Rate (%)"
  ) +
  theme_minimal()

Graph_Average_Lower_Middle_Income <- ggplot(Average_Lower_Middle_Income) +
  aes(x = Year, y = Mean_GDP_Growth_Rate, colour = Continent) +
  geom_line(linewidth = 0.2) +
  geom_hline(yintercept = 5, linetype = "dashed", colour = "black") +
  labs(
    title = "Average GDP per capita Growth Rate per Continent against Time (Lower Middle Income)",
    x = "Year",
    y = "GDP per capita Growth Rate (%)"
  ) +
  theme_minimal()

Graph_Average_Upper_Middle_Income <- ggplot(Average_Upper_Middle_Income) +
  aes(x = Year, y = Mean_GDP_Growth_Rate, colour = Continent) +
  geom_line(linewidth = 0.2) +
  geom_hline(yintercept = 4, linetype = "dashed", colour = "black") +
  labs(
    title = "Average GDP per capita Growth Rate per Continent against Time (Upper Middle Income)",
    x = "Year",
    y = "GDP per capita Growth Rate (%)"
  ) +
  theme_minimal()

Graph_Average_High_Income <- ggplot(Average_High_Income) +
  aes(x = Year, y = Mean_GDP_Growth_Rate, colour = Continent) +
  geom_line(linewidth = 0.2) +
  geom_hline(yintercept = 2, linetype = "dashed", colour = "black") +
  labs(
    title = "Average GDP per capita Growth Rate per Continent against Time (High Income)",
    x = "Year",
    y = "GDP per capita Growth Rate (%)"
  ) +
  theme_minimal()

Graph_Average_Low_Income
```

---

This chunk counts how many countries each continent has in total.
We will use this as the denominator when computing proportions
that meet their growth targets.

```{r total-countries-per-continent}
Total_countries_in_Continent <- Sorted_continents %>%
  group_by(Continent) %>%
  summarise(Number_Countries = n())
```

---

Here we calculate, for each income group, how many countries per
continent meet their respective growth target in each year
(7%, 5%, 4%, or 2%).

```{r countries-meeting-targets}
Low_Income_With_Growth <- Low_Income %>%
  filter(GDP_Growth_Rate >= 7) %>%
  group_by(Year, Continent) %>%
  summarise(Number = length(unique(Entity))) %>%
  ungroup()

Lower_Middle_Income_With_Growth <- Lower_Middle_Income %>%
  filter(GDP_Growth_Rate >= 5) %>%
  group_by(Year, Continent) %>%
  summarise(Number = length(unique(Entity))) %>%
  ungroup()

Upper_Middle_Income_With_Growth <- Upper_Middle_Income %>%
  filter(GDP_Growth_Rate >= 4) %>%
  group_by(Year, Continent) %>%
  summarise(Number = length(unique(Entity))) %>%
  ungroup()

High_Income_With_Growth <- High_Income %>%
  filter(GDP_Growth_Rate >= 2) %>%
  group_by(Year, Continent) %>%
  summarise(Number = length(unique(Entity))) %>%
  ungroup()

Achieved_Growth <- rbind(
  Low_Income_With_Growth,
  Lower_Middle_Income_With_Growth,
  Upper_Middle_Income_With_Growth,
  High_Income_With_Growth
)
```

---

We now convert these counts into proportions of countries
per continent that meet the GDP growth target each year,
and we visualise them with faceted bar charts.

```{r proportion-meeting-targets, fig.width=8, fig.height=5}
Proportion_Achieved_Growth <- Achieved_Growth %>%
  group_by(Year, Continent) %>%
  summarise(Total_Number = sum(Number)) %>%
  ungroup() %>%
  left_join(Total_countries_in_Continent, by = "Continent") %>%
  mutate(Proportion = Total_Number / Number_Countries)

Graph_Proportion_Achieved_Growth <- ggplot(Proportion_Achieved_Growth) +
  aes(x = Year, y = Proportion, fill = Continent) +
  geom_col() +
  facet_wrap(~Continent) +
  labs(
    x = "Year",
    y = "Proportion",
    title = "Proportion of Countries that met GDP per capita Growth Rate Target by Continent"
  ) +
  theme_minimal() +
  theme(strip.text = element_text(size = 15))

Graph_Proportion_Achieved_Growth
```

---

## Part 2 – Least Developed Countries (LDCs)

We import the UNSD methodology file that lists which countries are LDCs,
clean the table, and keep only the LDC rows.

```{r load-ldc-list}
Additional_csv_1 <- read.csv("~/Documents/GitHub/Data-Science-Coursework/UNSD — Methodology.csv",
                             header = FALSE, sep = ";")

colnames(Additional_csv_1) <- Additional_csv_1[1, ]

Additional_csv_1$"Least Developed Countries (LDC)" <-
  as.character(Additional_csv_1$"Least Developed Countries (LDC)")

Additional_csv_1$"Least Developed Countries (LDC)"[
  Additional_csv_1$"Least Developed Countries (LDC)" == ""
] <- NA

LDCs <- Additional_csv_1[-1, ] %>%
  select("Country or Area", "Least Developed Countries (LDC)") %>%
  filter(!is.na(`Least Developed Countries (LDC)`))
```

---

Here we merge the LDC list with the full GDP growth dataset,
obtaining GDP growth rates only for LDCs.

```{r ldc-merge-growth}
LDCs_GDP_Growth_Rate <- inner_join(
  LDCs,
  Countries_GDP_Growth_Rate,
  by = c("Country or Area" = "Entity")
)

LDCs_GDP_Growth_Rate <- LDCs_GDP_Growth_Rate[, -2]
```

---

This plot shows GDP per capita growth rates over time for all LDCs,
with a dashed 7% target line.

```{r ldc-growth-plot, fig.width=8, fig.height=5}
Graph_LDCs <- ggplot(LDCs_GDP_Growth_Rate) +
  aes(x = Year, y = GDP_Growth_Rate,
      group = `Country or Area`, colour = `Country or Area`) +
  geom_line(linewidth = 0.2) +
  geom_hline(yintercept = 7, linetype = "dashed",
             colour = "black", linewidth = 0.2) +
  labs(
    title = "GDP per capita Growth Rate against Time (LDCs)",
    x = "Year",
    y = "GDP per capita Growth Rate (%)",
    colour = "Country"
  ) +
  theme_minimal()

Graph_LDCs
```

---

We then restrict attention to the most recent 5 years for LDCs,
based on the latest year available in the data, and plot the
recent-period growth rates.

```{r ldc-recent-period, fig.width=8, fig.height=5}
max_year <- max(LDCs_GDP_Growth_Rate$Year, na.rm = TRUE)

LDC_GDP_Growth_Recent <- LDCs_GDP_Growth_Rate %>%
  filter(Year >= max_year - 4)

Graph_LDCs_Recent <- ggplot(LDC_GDP_Growth_Recent) +
  aes(x = Year, y = GDP_Growth_Rate,
      group = `Country or Area`, colour = `Country or Area`) +
  geom_line(linewidth = 0.2) +
  geom_hline(yintercept = 7, linetype = "dashed",
             colour = "black", linewidth = 0.2) +
  labs(
    title = "GDP per capita Growth Rate agianst Time (LDCs)",
    x = "Year",
    y = "GDP per capita Growth Rate (%)",
    colour = "Country"
  ) +
  theme_minimal()

Graph_LDCs_Recent
```

---

Here we compute average GDP growth rates by continent among LDCs,
and plot how these averages evolve over time with the 7% target line.

```{r ldc-continents-mean, fig.width=8, fig.height=5}
Continents_LDC_Growth_Rate <- LDCs_GDP_Growth_Rate %>%
  group_by(Continent, Year) %>%
  summarise(Mean_Growth = mean(GDP_Growth_Rate, na.rm = TRUE)) %>%
  ungroup()

Graph_Continents_LDC_Growth_Rate <- ggplot(Continents_LDC_Growth_Rate) +
  aes(x = Year, y = Mean_Growth, colour = Continent) +
  geom_line(linewidth = 0.2) +
  geom_hline(yintercept = 7, linetype = "dashed",
             colour = "black", linewidth = 0.2) +
  labs(
    title = "Average GDP per capita Growth Rate by Continent (LDCs)",
    x = "Year",
    y = "Average GDP per capita growth rate (%)",
    colour = "Continent"
  ) +
  theme_minimal()

Graph_Continents_LDC_Growth_Rate
```

---

Next we keep only year 2021 for LDCs and classify them into
growth‐rate categories, which will be used for a world map.

```{r ldc-2021-categories}
LDCs_GDP_Growth_Rate_2021 <- LDCs_GDP_Growth_Rate %>%
  filter(Year == 2021)

LDCs_GDP_Growth_Rate_2021 <- LDCs_GDP_Growth_Rate_2021 %>%
  mutate(Growth_Category = case_when(
    GDP_Growth_Rate >= 7  ~ "≥ 7% (Meets Target)",
    GDP_Growth_Rate >= 4  ~ "4% – 7%",
    GDP_Growth_Rate >= 0  ~ "0% – 4%",
    TRUE                  ~ "Negative Growth"
  ))
```

---

This chunk constructs a world map using `rnaturalearth`,
joins it with the LDC growth categories, and visualises the
distribution of LDC growth performance in 2021.

```{r ldc-world-map, fig.width=9, fig.height=5}
World <- ne_countries(scale = "medium", returnclass = "sf")

World_LDCs <- World %>%
  left_join(LDCs_GDP_Growth_Rate_2021, by = c("iso_a3" = "Code"))

Graph_World <- ggplot(World_LDCs) +
  geom_sf(aes(fill = Growth_Category), color = "grey", size = 0.1) +
  scale_fill_manual(
    values = c(
      "≥ 7% (Meets Target)" = "darkgreen",
      "4% – 7%"             = "lightgreen",
      "0% – 4%"             = "orange",
      "Negative Growth"     = "red"
    )
  ) +
  labs(
    title = "LDCs GDP per capita Growth Rate Categories (2021)",
    fill  = "Growth Category"
  ) +
  theme_minimal()

Graph_World
```

---

Finally, we look at how many LDCs **achieve or fail** the 7% target
over 2017–2021 by continent.
We first classify each observation by “Meets target” vs “Below target”,
then count the number of LDCs in each category.

```{r ldc-achievement-counts}
LDCs_According_Growth <- LDC_GDP_Growth_Recent %>%
  mutate(
    Target = ifelse(GDP_Growth_Rate >= 7,
                    "Meets target (≥ 7%)",
                    "Below target")
  ) %>%
  group_by(Year, Target, Continent) %>%
  summarise(LDCs_Number = length(unique(Code))) %>%
  ungroup()
```

---

We construct a full grid of all combinations of year, target status
and continent (to include zero counts), build a custom combined
`Year_Status` variable, and prepare labels for a two-row x-axis.

```{r ldc-achievement-grid}
All_Years      <- 2017:2021
All_Targets    <- c("Meets target (≥ 7%)", "Below target")
All_Continents <- unique(LDC_GDP_Growth_Recent$Continent)

Full_Grid <- expand.grid(
  Year      = All_Years,
  Target    = All_Targets,
  Continent = All_Continents,
  stringsAsFactors = FALSE
)

LDCs_According_Growth_Complete <- Full_Grid %>%
  left_join(LDCs_According_Growth,
            by = c("Year", "Target", "Continent")) %>%
  mutate(LDCs_Number = ifelse(is.na(LDCs_Number), 0, LDCs_Number))

LDCs_According_Growth_Complete <- LDCs_According_Growth_Complete %>%
  mutate(
    Growth_Status = ifelse(Target == "Meets target (≥ 7%)", "Achieved", "Did Not Achieve"),
    Year_Status   = paste0(Year, " (", Growth_Status, ")")
  )

Order <- LDCs_According_Growth_Complete %>%
  arrange(Year, desc(Growth_Status)) %>%
  pull(Year_Status) %>%
  unique()

LDCs_According_Growth_Complete$Year_Status <- factor(
  LDCs_According_Growth_Complete$Year_Status,
  levels = Order
)

x_labels <- LDCs_According_Growth_Complete %>%
  distinct(Year_Status, Year, Growth_Status)
```

---

This final graph displays, for each year, how many LDCs achieved
or did not achieve the 7% growth target, broken down by continent.
The x-axis uses two rows of labels: year and achievement status.

```{r ldc-achievement-plot, fig.width=9, fig.height=5}
Graph_LDCs_Growth_Status <- ggplot(LDCs_According_Growth_Complete) +
  aes(x = Year_Status, y = LDCs_Number, fill = Continent) +
  geom_bar(stat = "identity") +
  labs(
    x = "Year and Achievement Status",
    y = "Number of LDCs",
    title = "Number of LDCs that did not achieve or achieved 7% GDP per capita Growth Rate",
    fill = "Continent"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    plot.margin = margin(20, 20, 40, 20)
  ) +
  geom_text(aes(x = Year_Status, y = -0.5, label = Year),
            data = x_labels, inherit.aes = FALSE, vjust = 1.2) +
  geom_text(aes(x = Year_Status, y = -1.5, label = Growth_Status),
            data = x_labels, inherit.aes = FALSE, vjust = 1.2) +
  coord_cartesian(clip = "off")

Graph_LDCs_Growth_Status
```

